---
title: "Homework 6-1"
---

```{r}
library(dbplyr)
library(duckdb)
library(DBI)
library(tidyverse)
```

```{r}
conn <- dbConnect(duckdb(), dbdir = 'database.duckdb')
dbListTables(conn)
```

# 1. Claude

```{r}
dbGetQuery(conn, 
"SELECT p.Name, COUNT(*) AS num_nests
FROM Bird_nests bn
JOIN Personnel p ON bn.Observer = p.Abbreviation
WHERE bn.Site = 'nome'
  AND bn.Year BETWEEN 1998 AND 2008
  AND bn.ageMethod = 'float'
GROUP BY p.Name
HAVING COUNT(*) = 36;")
```

This query generated by Claude looks pretty standard and does show a good understanding of the database. The result ended up being correct.

# 2. ChatGPT

```{r}
dbGetQuery(conn, 
"SELECT
    p.Name AS observer_name,
    COUNT(*) AS num_nests
FROM Bird_nests b
JOIN Personnel p
    ON b.Observer = p.Abbreviation
WHERE b.Site = 'nome'
  AND b.Year BETWEEN 1998 AND 2008
  AND b.ageMethod = 'float'
GROUP BY p.Name
HAVING COUNT(*) = 36;")
```

This query was generated by ChatGPT and is pretty much the same as Claude's response. The result also is correct.

#3. Gemini

```{r}
dbGetQuery(conn,
"SELECT 
    p.Name, 
    COUNT(bn.Nest_ID) AS num_nests
FROM Bird_nests bn
JOIN Personnel p ON bn.Observer = p.Abbreviation
JOIN Site s ON bn.Site = s.Code
WHERE 
    LOWER(s.Site_name) = 'nome' 
    AND bn.Year BETWEEN 1998 AND 2008
    AND bn.ageMethod = 'float'
GROUP BY p.Name
HAVING COUNT(bn.Nest_ID) = 36;")
```

This query is the most different compared to the last two. While the result was correct, it used COUNT(Nest_ID), the exact misunderstanding that Nest_ID can't be NULL and the nest table.