---
title: "Data Cleaning Assignment"
---

```{r}
library(tidyverse)
library(here)

# import data
snowsurvey <- read_csv(here("data", "processed", "snow_cover.csv")) |> 
  janitor::clean_names()
```

```{r}
# convert string values to numeric, replace non-num values to NA
snowsurvey_num <- snowsurvey |> 
  mutate(across(c(water_cover, land_cover, total_cover), ~ as.numeric(.x)))
```

String values handled, need to handle values outside of range 0-100

```{r}
# if values are less than 0 or greater than 100, assign to NA
snowsurvey_range <- snowsurvey_num |> 
  mutate(across(c(water_cover, land_cover), 
  ~ ifelse(.x < 0 | .x > 100, NA, .x)))
```

First, some rows are all NA, with a couple cases:

1. [NA, NA, NA, NA]
2. [NA, NA, NA, REAL], where REAL is an actual number in the total_cover columns

both cases are not useful and should be dropped.

```{r}
# drop rows where snow, water, and land cover columns are ALL NA
snowsurvey_fixed <- snowsurvey_range |> 
  mutate(all_na = rowSums(is.na(across(c(snow_cover, water_cover, land_cover)))) == 3) |> 
  filter(all_na == FALSE)
```

Now that rows with all NAs are dropped, we still have rows with 1 NA that should be zero, in rows like this: [NA, 90, 10, 100]

  - These will be imputed.

```{r}
# impute rows with 1 NA
snowsurvey_fixed <- snowsurvey_fixed |> 
  mutate(across(c(snow_cover, water_cover, land_cover), 
  ~ if_else(
    rowSums(is.na(across(c(snow_cover, water_cover, land_cover)))) == 1 & is.na(.x), # if TRUE
    100 - rowSums(across(c(snow_cover, water_cover, land_cover)), na.rm = TRUE), # impute
    .x
  )))
```

We now are left with rows with two NAs:

1. Imputable case [NA, NA, 100, 100], or a similar combination, fill NAs with zeros
2. Impossible imputable case [45, NA, NA, 100], these are impossible to impute, we can drop these rows.
3. Invalid case [50, NA, NA, 50], since we defined the total column to only be 100, we can drop these rows too.

```{r}
# case 1, impute two na case
snowsurvey_fixed <- snowsurvey_fixed |> 
  mutate(across(c(snow_cover, water_cover, land_cover), 
  ~ if_else(
    rowSums(is.na(across(c(snow_cover, water_cover, land_cover)))) == 2 & 
    rowSums(across(c(snow_cover, water_cover, land_cover)), na.rm = TRUE) == 100 &
    is.na(.x),
    0,  # if one value is 100, the other two must be 0
    .x
  )))

# case 2/3
snowsurvey_fixed <- snowsurvey_fixed |> 
  filter(!(total_cover != 100 & 
           rowSums(is.na(across(c(snow_cover, water_cover, land_cover)))) > 0))
```

There still seem to be a few cases to handle:

1. *_cover columns do correctly add to 100, total_cover is 100. [90, 0, 10, 100], keep these rows.
2. *_cover columns don't add up to 100, total_cover is 100. [0, 0, 0, 100], drop these rows.
3. *_cover columns don't add up to 100, total_cover is not 100. [0, 0, 15, 85], drop these rows.


```{r}
# only keep rows that add to 100
snowsurvey_fixed <- snowsurvey_fixed |> 
  filter(
    snow_cover + water_cover + land_cover == 100,
    total_cover == 100
  )
```

The final output is a dataframe with correct snow, water, & land cover columns which add to 100.

```{r}
# write to csv
write_csv(snowsurvey_fixed, "data/processed/all_cover_fixed_zach.csv")
```